\section{Extended Logic Counting Classes}

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
	\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
	&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
	\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. For every logic $\mathcal{F}$, we define an $\mathcal{F}$-extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
	\varphi &::=& \alpha, \alpha \text{ is an $\mathcal{F}$-formula over } \L  \ \mid \\
	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
	f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.\\

\subsection{The First {\sc FO}-Extended Counting Hierarchy}

We define syntactically the classes $\XE{i}$ and $\XU{i}$ according to the following grammar:
\begin{enumerate}
	\item $\XE{0}$:
	\[
		E_0 ::= \varphi , \varphi \mbox{ is an {\sc FO}-extended quantifier-free $\L$-formula }.
	\]
	\item $\XU{0}$:
		\[U_0 ::= E_0.\]
	\item $\XE{i+1}$:
	\[	E_{i+1} ::= U_i \ \mid \ \exists x \, E_{i+1}.\]
	\item $\XU{i+1}$:
		\[U_{i+1} ::= E_i \ \mid \ \forall x \, U_{i+1}.\]
\end{enumerate}

A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.

\begin{theo} \label{fp1}
	$\XE{0} \subseteq$ {\sc FP}.
\end{theo}
\begin{proof}
	Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
	\begin{eqnarray*}
		f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
	\end{eqnarray*}
	for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates. We will now show that computing $f(\A)$ can be done in polynomial time.
	
	For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$. Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \Truc$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$. Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time. Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.
	
	Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$. Also, let $g = f_{\psi(\X,\x)}$. Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$. But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$. Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.
	
\end{proof}

The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \Truc \mid f(\A) > 0\}$.

\begin{theo} \label{decisionptime}
	The decision problem associated to a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
	Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
	\end{eqnarray*}
	where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$. Moreover, let $g = f_{\psi(\X,\y)}$.
	\begin{claim} \label{mono}
		For each $\A \in \Truc$, $f(\A) > 0$ iff $g(\A) > 0$.
	\end{claim}
	\begin{proof}
		($\Rightarrow$) Suppose $f(\A) > 0$. Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$. It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$. Therefore, $g(\A) > 0$. 
		
		($\Leftarrow$) Suppose $g(\A) > 0$. Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$. Then we have that $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
	\end{proof}
	Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\XE{0}$. It follows from \ref{mono} that for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$. But by Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time. From this we conclude that the decision version of $f$ is in $\ptime$.
\end{proof}

For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
\begin{eqnarray*}
	(f \dotminus g)(\A) =
	\begin{cases}
		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
		0, & \text{if }f(\A) \leq g(\A).
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.

\begin{theo}
	If $\E1$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theo}
\begin{proof}
	Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there exists $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.
	
	Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
	\begin{eqnarray*}
		S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
		S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
		S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
		S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
	\end{eqnarray*}
	Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
	\begin{multline*}
	\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
	\end{multline*}
	Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
	$$
	\varphi(x,X) = (T(x) \vee \neg T(x)).
	$$
	Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. Given that $f_{\#3DNF}, f_{all} \in \E1$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E1$. Let $h\in\E1$ be such that $h = f_{all}-f_{\#3DNF}$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete. However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E1$, we have that $L_h \in \ptime$. Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.
\end{proof}

\begin{theo}
	$\#\Sigma_1 \subsetneq \E1$
\end{theo}
\begin{proof}
	We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Note that $\A$ is an induced substructure of $\B$.
	
	We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\Truc$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}. Therefore, $f(\B) \geq f(\A) = 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
\end{proof}

For a given function $f$, we define $f \dotminus 1$ as follows:
\begin{eqnarray*}
	f \dotminus 1(\A) =
	\begin{cases}
		f(\A)-1, & \text{if }f(\A) > 0 \\
		0, & \text{if }f(\A) = 0.
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.

\input{th4}

\begin{theo}
	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
\end{theo}
\begin{proof}
	Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}. We will now show a function $h:\Truc\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time. For every DNF formula $\Phi$ we define $\A_{\Phi} \in \textsc{Struct}[\L_{\textsc{DNF}}]$ as stated in \ref{dnfdef}. Let $\psi$ be such that $f = f_{\psi(\X,\z)}$. We separate the proof in three cases:
	\begin{enumerate}
		\item $\X = ()$. Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = h(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The DNF formula we generate is
		\[
		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
		\]
		which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
		
		\item $\z = ()$. Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = h(\A)$ as follows. First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$. Let each $\varphi_{i}(\x,\X)$ be of the form
		\[
		\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
		\]
		where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula. We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.
		
		We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$. Let $\p$ be the tuple of all such variables.
		
		Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$. There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
		
		Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
		\[
		\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
		\]
		and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
		
		As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$. We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$. We generate the tuple of variables $\p = (p^1_1,p^1_2,p^1_3) \cup (p^2_{(i,j)})$ for $(i,j)\in \{1,2,3\}\times\{1,2,3\}$. Note that in this procedure we will only have one set $\Gamma_1 = \{(1,2),(2,3)\}$. And so, the DNF formula we generate is 
		\begin{multline*}
		\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
		\end{multline*}
		
		Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$. We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
		
		\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = h(\A)$ as follows. As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$. Let each $\varphi_{i}(\x,\X,\z)$ be of the form
		\[
		\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
		\]
		where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula. We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.
		
		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$. This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$. Let $\p$ be the tuple of all such variables. We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$. Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
		
		For each $\e\in A^d$, we follow the following procedure. For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
		\[
		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
		\]
		where $\gamma_{\e}$ is defined as
		\[
			\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
		\]
		and $\theta_{\e}$ is defined as
		\[
			\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
		\]
		Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
		
		As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$. We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$. We generate the tuple of variables $\p = (p^e_{1,1},p^e_{1,2},p^e_{1,3},p^e_{2,1},p^e_{2,2},p^e_{2,3})$, for $e\in \{1,2,3\}$. Here, we consider $L = 2$, $T = \{t_1,t_2\}$, and $\beta$ defined as $\beta(1) = (0,0)$, $\beta(2) = (0,1)$ and $\beta(3) = (1,0)$. In this procedure we generate the sets $\Gamma^1_1 = \emptyset$, $\Gamma^2_1 = \{1\}$, $\Gamma^3_1 = \{1,2\}$. Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
		\[
		\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
		\]
		and $\theta_e$ for $e\in\{1,2,3\}$,
		\begin{align*}
		\theta_1 &= \neg t_1 \wedge \neg t_2,\\
		\theta_2 &= \neg t_1 \wedge t_2,\\
		\theta_3 &= t_1 \wedge \neg t_2.		
		\end{align*}
		Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
		\begin{align*}
		\Phi^1_1 &= \square\\
		\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
		\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
		\end{align*}
		and so, the DNF formula we generate is
		\begin{multline*}
		\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
		(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
		(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
		(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
		(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
		(p^e_{2,3}\wedge\neg p^e_{2,3})].
		\end{multline*}
		To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection. Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows. For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$. For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$. For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$. 
		
		Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$. It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$. Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$. Therefore, $g$ is an injective function. Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$. Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$. Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$. Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$. Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$. From this we deduce that $g$ is surjective, and therefore a bijection. We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
	\end{enumerate}
\end{proof}
\begin{theo}
	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
\end{theo}
\begin{proof}
	Suppose it is not properly contained for $n$. Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$. Note that $f_{\varphi(\z)}(\A) = 1$ if $\A$ has at least $n+1$ elements, and  $f_{\varphi(\z)}(\A) = 0$ otherwise, for each $\A\in\Truc$. Let $\forall\x\,\psi(\x,\X,\u)$ be such that $f_{\varphi(\z)} = f_{\forall\x\,\psi(\x,\X,\u)}$. Let $\A = \langle A, \S^{\A}, <^{\A} \rangle\in\Truc$ where $\vert A \vert = n+1$. Note that $f_{\forall\x\,\psi(\x,\X,\u)}(\A) = 1$. Let $(\P,\e)$ be such that $\A\models\forall\x\,\psi(\x,\P,\e)$. Let $a\in A$ be such that $a\not\in\e$. Consider the vocabulary $\L' = \L \cup \{\Q\}$, and let $\A' = \A \cup \langle \Q^{\A'} \rangle$ where $\Q^{\A'} = \P$. Note that $\A\models\varphi(\P,\e)$ if and only if $\A'\models\varphi(\e)$. Consider $\B'$ as the induced substructure of $\A'$ with domain $A\setminus\{a\}$, and let $\Q^{\B'}$ be the induced interpretation of $\Q$. Also let $\B$ be the induced substructure of $\A$ with domain $A\setminus\{a\}$. As universal formulae are preserved under substructure\footnote{Hodges, 1997. P. 183}, we have that $\B'\models\forall\x\,\psi(\x,\e)$, which implies that $\B\models\forall\x\,\psi(\x,\Q^{\B'},\e)$. We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
\end{proof}