\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm} 
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}

\def\A{{\frak A}}

\newtheorem{theo}{Theorem}
\newtheorem{lemma}[theo]{Lemma}
\newtheorem{claim}[theo]{Claim}
\newtheorem{coro}[theo]{Corollary}

\begin{document}

\begin{flushleft}
{\footnotesize Pontificia Universidad Cat\'olica de Chile - Facultad de Ingenier\'ia\\
Departamento de Ciencias de la Computaci\'on\\
}
\end{flushleft}
\begin{center}
{ \LARGE \bf
  Some properties of $\#\Sigma_1^{+}$
}
\end{center}

The $\#\Sigma_1^{+}$ class is defined as an extension of $\#\Sigma_1$ which includes the possibility of adding arbitrarily quantified sub-formulas that don't include counted predicates.

Formally, let $\vec{P}$ be a sequence of predicates and $\vec{x},\vec{z}$ sequences of variables. Let $\varphi(\vec{S},\vec{x},\vec{z})$ be a formula where the variables in $\vec{P}$ are not in a quantified sub-formula. Note that sub-formulas that include any other predicate are not restricted at all. A counting problem $F \in \#\Sigma_1^{+}$ if given a structure $\A$
$$
F(\A) = \vert \{ \langle\vec{P},\vec{z}\rangle :\A \models \exists \vec{x} \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
$$ 
Some of its properties are:

\begin{theo}
The decision version of a counting problem in $\#\Sigma_1^{+}$ is in \text{P}.
\end{theo}
\begin{proof}
Let $F$ be a function in $\#\Sigma_1^{+}$. Then there is a formula $\varphi$ such that:
$$
F(\A) = \vert \{ \langle\vec{P},\vec{z}\rangle :\A \models \exists \vec{x} \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
$$ 
where $\A$ is a finite ordered structure and $\vec{z}$ is an $m$-tuple. For each $\vec{z} \in A^m$, every arbitrarily quantified sub-formula which does not include $\vec{P}$ can be evaluated in polynomial time. Let $\varphi'$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula by a contradiction. Note that $\varphi'$ is quantifier-free.

Then, we compute a similar counting function:
$$
F'(\A) = \vert \{ \langle\vec{P},(\vec{z},\vec{x})\rangle :\A \models \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
$$ 
which is in $\#\Sigma_0$. Note that for every $\A$, $F(\A) > 0$ iff $F'(\A) > 0$, so computing $F'$ is enough to solve the decision version of $F$. However, Saluja and Subrahmanyam showed that any counting problem in $\#\Sigma_0$ is computable in polynomial time \footnote[1]{(paper)}, therefore, the decision version of $F$ is in P.
\end{proof}

\begin{theo}
$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
\end{theo}
\begin{proof}
\#3DNF $\in \#\Sigma_1$. Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance. Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$. This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete. However, as we showed previously (Theorem 1), it's also in P. Then, co-NP $\subseteq$ P.
\end{proof}

\begin{theo}
$\#\Sigma_1 \subsetneq \#\Sigma_1^{+}$
\end{theo}
\begin{proof}
We will show that the $\#\Sigma_1^{+}$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall x_2 S(x_2)$ is not in $\#\Sigma_1$.
\end{proof}

\begin{theo}
$\#\Sigma_1$ is closed under substraction of 1 element.
\end{theo}
\begin{proof}
The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
\begin{enumerate}
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{z}) \} \vert
$$
our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $F'$ be:
$$
F'(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
$$
where $\varphi_{\lneq}(\vec{z'},\vec{z})$ is a formula that indicates that $\vec{z'}$ is lexicographically smaller than $\vec{z}$. Thus, $F'(\A)$ will count exactly one element less than $F(\A)$ if $F(\A)>0$.
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{P}) \} \vert
$$
similarly, we would like to eliminate the lexicographically smallest predicate $P$ that holds the formula true with the exact number of tuples. First we treat sub-formulas without $\vec{P}$ as literals, and take a DNF formula that's equivalent to $\varphi$. Then
$$
F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
\exists \vec{x} \varphi_1(\vec{x},\vec{P}) \vee
\exists \vec{x} \varphi_2(\vec{x},\vec{P}) \vee
... \vee
\exists \vec{x} \varphi_k(\vec{x},\vec{P})
 \} \vert
$$
We can also suppose that all of this formulas are satisfied by a single total order for $\vec{x}$. We can manually eliminate unsatisfiable sub-formulas such as
$$
\exists x,y(P(x,y) \wedge \neg P(x,x) \wedge x = y)
$$
Therefore,
$$
\varphi_1(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_j(\vec{x_j}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})
$$
where $P_i$ and $\overline{P}$ are subsequences of $\vec{P}$, $P_i$ is of arity $a_i$ and $\vec{x_i}$ is an $a_i$-tuple of variables in $\vec{x}$, $\vec{y}$ is also a subsequence of $\vec{x}$, $\theta(\vec{x})$ defines a total order for $\vec{x}$, and $\varphi_{\vec{S}}(\vec{x})$ involves only closed predicates. Without unsatisfiable formulas, we can assure that if there is an assignation $\vec{x}$ that satisfies $\alpha(\vec{x}) = \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})$, there is at least one possible assignation to $\vec{P}$. We now define
\begin{multline*}
\alpha_{\min}^i(\vec{x_1}',...,\vec{x_n}') = \\ \exists\vec{y}\left( \alpha(\vec{x_1}',...,\vec{x_n}',\vec{y})\wedge \forall\vec{x_1}''...\vec{x_n}''\vec{y}'(\alpha(\vec{x_1}'',...,\vec{x_n}'',\vec{y}')\rightarrow (\vec{x_1}',...,\vec{x_n}') \leq (\vec{x_1}'',...,\vec{x_n}'')\right)
\end{multline}
Note that $\alpha_{\min}^i$ is satisfied only by the lexicographically smallest assignation $\vec{x}^*$ to $\vec{x}$ that satisfies both $\theta(\vec{x})$ and $\varphi_{\vec{S}}(\vec{x})$. Our new formula is
\begin{multline*}
\varphi_1'(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})\wedge\exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
\bigg(\bigvee_{P\in\vec{P}}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\vec{P}} \exists \vec{x}''\Big( P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
\end{multline}
where $\vec{x_i}'$ is a an $a_i$-tuple of variables in $\vec{x}$. If there is an assignation to satisfy $\alpha^j(\vec{x})$, this formula will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$. Let $\vec{P}^*$ be the excluded assignation to $\vec{P}$. Suppose there is another assignation $\vec{P}^*'$ to $\vec{P}$ that is also excluded by the new sub-formula. Then, $\vec{x}^*$ is such that every tuple $\vec{x}_i^*$ is in $\vec{P}_i^*'$ and $\vec{x}^*$ includes every tuple in $\vec{P}^*'$. If that's the case, $\vec{P}^*'$ contains exactly the same tuples as $\vec{P}^*$, and are thus the same. Now we need to exclude that assignation to the next disjuncts, which we do by adding the same sub-formula. We also need to include the case where there are no possible assignations to $\vec{x}$ for the first disjunct, so we add this sub-formula:
\begin{multline*}
\forall\vec{x}\left(\neg\alpha^1_{\min}(\vec{x})\right) \to \exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha^2_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
\bigg(\bigvee_{P\in\vec{P}}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\vec{P}} \exists \vec{x}''\Big( P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
\end{multline}
The full disjunct is as follows:
\begin{multline*}
\varphi_j'(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x}) \wedge \Bigg(\Big(\forall\vec{x}\left(\neg\alpha^1_{\min}(\vec{x})\right)\wedge ... \wedge \forall\vec{x}\left(\neg\alpha^{j-1}_{\min}(\vec{x})\right)\Big) \to \\ 
\exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha^j_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge
\bigg(\bigvee_{P\in\vec{P}}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\vec{P}} \exists \vec{x}''\Big( P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\Big) \bigg) \Bigg) \Bigg)
\end{multline}
which, if all of the previous disjuncts have no possible assignations, eliminates the least assignation to $\vec{P}$ in that disjunct. Finally, the counting set
$$
F'(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
\exists \vec{x} \varphi'_1(\vec{x},\vec{P}) \vee
\exists \vec{x} \varphi'_2(\vec{x},\vec{P}) \vee
... \vee
\exists \vec{x} \varphi'_k(\vec{x},\vec{P})
 \} \vert
$$
counts exactly one assignation to $\vec{P}$ less than $F(\A)$.

\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{P},\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x}, \vec{P},\vec{z}) \} \vert
$$
Then, we going to isolate the minimal predicate $\vec{P}$ that holds the formula true and eliminate the lexicographically smallest $\vec{z}$ that satisfies it. We mix both previous strategies.
\end{enumerate}
\end{proof}

\end{document}
